---
title: "GGPlot Continued"
author: "John W. Hoggard"
date: "8/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Basic Scatterplots and Smoothers

Set up a ggplot, add data, aesthetic mappings, and then finally a geom:

```{r}
library(ggplot2)
ggplot(data = iris) +
    geom_point(aes(x = Sepal.Length,
                   y = Sepal.Width))
```

We can add a mapping to identify species by color:

```{r}
ggplot(data = iris) +
    geom_point(aes(x = Sepal.Length,
                   y = Sepal.Width, 
                   color = Species))
```

```{r}
str(iris)
```

Remember in base R plotting, we had to extract columns from the data frame:
```{r}
plot(iris$Sepal.Length, iris$Sepal.Width)
```

Could also map, say, shape:
```{r}
ggplot(data = iris) +
    geom_point(aes(x = Sepal.Length,
                   y = Sepal.Width, 
                   shape = Species))
```

```{r}
ggplot(data = iris) +
    geom_point(aes(x = Sepal.Length,
                   y = Sepal.Width, 
                   color = Species)) +
    geom_smooth(aes(x = Sepal.Length, y = Sepal.Width), 
                method = "loess", se = FALSE)
```

```{r}
ggplot(data = iris) +
    geom_point(aes(x = Sepal.Length,
                   y = Sepal.Width, 
                   color = Species)) +
    geom_smooth(formula = y ~x, aes(x = Sepal.Length, y = Sepal.Width), 
                method = "lm", se = FALSE)
```


```{r}
ggplot(data = iris,
       aes(x = Sepal.Length,
                   y = Sepal.Width)) +
    geom_point(aes(color = Species)) +
    geom_smooth(formula = y ~x, method = "lm", se = FALSE,
                aes(group = Species))
```

Or we can set the colors for everything:
```{r}
ggplot(data = iris,
       aes(
           x = Sepal.Length, 
           y = Sepal.Width,
           color = Species
       )
       ) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE, formula = y~x)
```

An alternate way to look at subsets is to *facet* the plot:
```{r}
ggplot(data = iris, aes(x = Sepal.Length, y = Sepal.Width)) +
    geom_point() +
    geom_smooth(formula = y~x, method = "lm", se = FALSE) +
    facet_wrap( ~ Species, nrow = 3)
```

## Some Other Geoms

There are tons of geoms available.  (Type `geom` and use autocomplete to check.)

Histogram of Sepal.Length:
```{r}
ggplot(data = iris) + 
    geom_histogram(aes(x = Sepal.Length), binwidth = .5)
```

(Set bins or binwidth to adjust the number of bins.)

What if we set a fill color? Outside of the aesthetic:
```{r}
ggplot(data = iris) + 
    geom_histogram(aes(x = Sepal.Length), binwidth = .5, 
                   fill="blue", 
                   color = "green")
```

Inside the aesthetic by mapping to Species, for example:
```{r}
ggplot(data = iris) +
    geom_histogram(aes(x = Sepal.Length, fill = Species), binwidth = .5)
```

These are plotted on top of each other:
```{r}
ggplot(data = iris) +
    geom_histogram(aes(x = Sepal.Length, fill = Species), binwidth = .5) +
    facet_wrap(~Species)
```

You could also use boxplots:
```{r}
ggplot(data = iris) +
    geom_boxplot(aes(x = Species, y = Sepal.Length))
```

## Adding Style to the Graph

Add on labels with the `labs` function:
```{r}
ggplot(data = iris) +
    geom_boxplot(aes(x = Species, y = Sepal.Length)) +
    labs(title = "Distribution of Sepal Lengths",
         y = "The Sepal Length")
```

Something cool:
```{r}
ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width)) +
    geom_point(color = "white", size = 3) +
    geom_point(aes(color = Species)) + 
    labs(title = "Cool looking points",
         x = "Sepal Length", 
         y = "Sepal Width")
```

