---
title: "More Tidying"
author: "John W. Hoggard"
date: "10/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Back to tidyr

```{r}
library(tidyverse)
library(here)
```

Recall:
```{r}
table1
```

We can `pivot_wider` to switch to a wide format (the `year` column really contains variables):
```{r}
pivot_wider(table1,
            names_from = year,
            values_from = cases:population)
```

Or we can switch to a longer format:
```{r}
pivot_longer(
    table1,
    cases:population,
    names_to = "Type",
    values_to = "Value"
)
```

## Tutoring Data

Tutoring log:
```{r}
tutor_logs <- read_csv(
    here("Datasets", "Tidying", "Tutoring.csv"),
    col_types = cols(
  Date = col_date(format = "%m/%d/%y"),
  Tutor = col_character(),
  Class = col_character(),
  `Time (in hours)` = col_double()
)
)
```

Fix the annoying name:
```{r}
tutor_logs %>%
    rename(Time = `Time (in hours)`) -> 
    tutor_logs
```


How about a report of hours by tutor and class?
```{r}
tutor_logs %>%
    group_by(Tutor, Class) %>%
    summarize(total_time = sum(Time))
```

Maybe this would be better pivoted wide?

```{r}
tutor_logs %>%
    group_by(Tutor, Class) %>%
    summarize(total_time = sum(Time)) %>%
    pivot_wider(
        names_from = Class,
        values_from = total_time,
    )
```

If we want total hours taught per tutor, that's easy enough too:
```{r}
tutor_logs %>%
    group_by(Tutor) %>%
    summarize(total_time = sum(Time))
```

Or by class:
```{r}
tutor_logs %>%
    group_by(Class) %>%
    summarize(total_time = sum(Time))
```

## Separating and Uniting

Sometimes data in a column is really more than just a single variable, and we have to split a column into multiple variables.
```{r}
math_peeps <- tribble(
    ~Name, ~Dates,
    #-----|--------
    "Pierre de Fermat", "1607-1665",
    "Augustin-Louis Cauchy", "1789-1857",
    "Bernhard Riemann", "1826-1866"
)
```

Included in `tidyr` is `separate`, which will split a column into 2 (or more) columns.  By default, it will separate on any non-alphanumeric characters. We have to name the new columns with the `into` argument:
```{r}
math_peeps %>%
    separate(Dates, 
             into = c("Born", "Died"))
```

```{r}
math_peeps2 <- tribble(
    ~Name, ~Dates,
    #-----|--------
    "Pierre de Fermat", "(1607-1665)",
    "Augustin-Louis Cauchy", "(1789-1857)",
    "Bernhard Riemann", "(1826-1866)"
)
```

We split on every non-alphanumeric character:
```{r}
math_peeps2 %>%
    separate(Dates, 
             into = c(NA, "Born", "Died", NA))
```

Shortcut if there are columns you just want to discard:  Name those columns with NAs.

The reverse:  Sometimes you want to paste stuff together.  Then we use `unite`:
```{r}
math_peeps %>%
    separate(Dates, 
             into = c("Born", "Died")) %>%
    unite("Lifespan", Born, Died, sep = " until ")
```

By default, unites with an underscore--or you can specify with `sep` argument.

There is a more general way to split with separate using a regular expression, and we'll cover these a bit later.

## Using Pivots for Graphing















