---
title: "Iteration"
author: "John W. Hoggard"
date: "11/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## For Loop

General structure:
```{r, eval = F}
for (val in vector) {
    <commands>
}
```

This is a "ranged" for loop; there is no incrementing a variable.
```{r}
wds <- c("Mary", "mary", "quite", "contrary")
for (x in wds) {
    print(x)
}
```

```{r}
library(stringr)
for (x in wds) {
    x <- str_to_upper(x)
    print(x)
}
```

But we didn't actually change any elements in the vector:
```{r}
wds
```

If I really want to make changes to the original vector, I could use an index, and access the elements directly, one at a time:
```{r}
for ( index in 1:4 ) {
    wds[index] <- str_to_upper(wds[index])
}
wds
```

Better:  Use `seq_along` to produce the indices we need:
```{r}
seq_along(wds)
```

Creates a vector of same length as the vector we start with.  Caution: It looks like the following would also work:
```{r}
1:length(wds)
```

But what about an empty vector:
```{r}
empty_vector <- vector(mode = "character", 0)
seq_along(empty_vector)
```

```{r}
1:length(empty_vector)
```

Let's generate some lists of random normal variables:
```{r}
ten_samples = list()

for (i in 1:10) {
    ten_samples[[i]] <- rnorm(20)
    # 20 values from random normal mean 0, sd = 1
}
```

```{r}
ten_samples
```

What are the means of these ten random samples?

A naive way to do this:
```{r}
avgs <- vector()
for ( numbers in ten_samples ) {
    avgs <- c(avgs, mean(numbers))
}
```

```{r}
avgs
```

This is inefficient. Each time we grow a vector, we have to reallocate memory space to hold that vector.  

Better: Preallocate the entire vector for the results first:
```{r}
avgs <- vector(mode = "double", 10)
for ( i in seq_along(ten_samples) ) {
    avgs[[i]] <- mean(ten_samples[[i]])
}
```

```{r}
avgs
```

Moral: Never "grow" a vector by adding to it in a loop.





