---
title: "Vectors"
author: "John W. Hoggard"
date: "11/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Vectors

Vectors are what we usually use for listings of alike items:
```{r}
x <- c(2, 3, 4, 7)
```

```{r}
x
```

In fact, everything is a vector:
```{r}
7
```

We have types:
```{r}
typeof(x)
```

There are also actually integers:
```{r}
typeof(1:5)
```

```{r}
typeof(c(1, 2, 3, 4, 5))
```
Also character types:
```{r}
typeof(c("Joe", "Sue"))
```

When you start combining items in a vector, it automatically switches to a common type:
```{r}
typeof(c(1:5, 2.0))
```

```{r}
typeof(c(1:5, "six"))
```

We can extract elements using `[]`:
```{r}
x[2:3]
```

If you want to extract exactly *one* thing from a vector, you can use `[[]]`:
```{r}
x[[2]]
```

Objects in R have attributes:
```{r}
attributes(x)
```

Named vectors have attributes for the name:
```{r}
y <- c("First" = 1, "Second" = 2)
y
```

```{r}
y[["Second"]]
```

```{r}
attributes(y)
```

Can get to the names with the helper function `names`:
```{r}
names(y) <- c("Shirley", "Fred")
y
```

Matrices have dimension attribute:
```{r}
mtx <- matrix(1:6, nrow = 2)
mtx
```

```{r}
attributes(mtx)
```

```{r}
dim(mtx) <- c(3, 2)
mtx
```

## Lists

Lists are kind of like vectors, but you can hold *anything*, and the elements can be of different types:
```{r}
z <- list("first" = 5, "second" = "Bob Smith", 
          "third" = 1:10)
z
```

I mean, we can can *really* put anything in a list:
```{r}
z2 <- list(5:10, z, c("a", "b", "c"))
z2
```

Extracting from a list can use `[]`, and can extract ranges--but I get another list:
```{r}
str(z[2:3])
```

```{r}
z[2]
```

But `[[]]` extracts the element:
```{r}
z[[2]]
```

Double brackets breaks out of list:
```{r}
z2[[2]]
```

Single brackets leaves the top level list intact:
```{r}
z2[2]
```

Can also index multiple times:
```{r}
z2[[2]][[3]]
```

We can also extract named elements from a list with `$`:
```{r}
z$first
```

```{r}
z$second
```

What's a data frame?
```{r}
df <- data.frame(x = 1:10, y = 2*(1:10))
str(df)
```

```{r}
df[[2]]
```

In other words, a data from is a list of vectors--with attributes.
```{r}
attributes(df)
```












