---
title: "Reading Data and EDA"
author: "John W. Hoggard"
date: "9/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reading Data with `readr`

There is a package called `readr` in the Tidyverse for reading data.

```{r}
library(tidyverse)
```

Get the pathname to the file with `here`:
```{r}
library(here)
simple_data_path <- here("Datasets", "SimpleData.csv")
```

Read it in with `readr`:
```{r}
simple_data <- read_csv(simple_data_path)
simple_data_base <- read.csv(simple_data_path)
```

```{r}
str(simple_data_base)
```

```{r}
str(simple_data)
```

We don't get munged column names, but it can be tricky to grab those columns:
```{r}
# Base R version:
simple_data_base$Employee.Name
```

ReadR version:
```{r}
simple_data$`Employee Name`
```

What is team exactly?
```{r}
simple_data %>%
    count(Team)
```

## Parsing

ReadR comes with functions which "parse" or read various types of data.  
For example, we can parse numerical data:
```{r}
x <- c("123", "4.53", "-99", "", "pi", "-4.78", "Cost is 3.50", "$128.34")
```

Use `parse_number(x)1 to parse a vector of strings into numbers:
```{r}
parse_number(x)
```

Now: When we use `read_csv`, it produces column specifications using related functions like `col_number()`.  Let's re-read the data, but change the Pledge column to be parsed as a number:

```{r}
simple_data <- read_csv(simple_data_path,
                        col_types = 
                            cols(
  `Employee Name` = col_character(),
  `ID #` = col_double(),
  `Phone Number` = col_character(),
  Address = col_character(),
  Email = col_character(),
  Pledge = col_number(),
  Team = col_character(),
  `Date Hired` = col_character()
)
)
```

```{r}
str(simple_data)
```

Of course can't do that if it's not a number:
```{r}
read_csv(simple_data_path,
                        col_types = 
                            cols(
  `Employee Name` = col_number(),
  `ID #` = col_double(),
  `Phone Number` = col_character(),
  Address = col_character(),
  Email = col_character(),
  Pledge = col_number(),
  Team = col_character(),
  `Date Hired` = col_character()
)
)
```

If you want to specifically have doubles, there is a `parse_double` (and `col_double`):
```{r}
parse_double(x)
```

Dates are a little tricky:
```{r}
parse_date(c("2020-09-15", "September 9, 2020", "9/15/20"))
```

We can get other things represented:
```{r}
parse_date(c("2020-09-15", "September 9, 2020", "9/15/20"),
           format="%m/%d/%y")
```

Read one last time, but with: Character for ID, factor for Team, dates for Date Hired, and numbers for Pledges:
```{r}
simple_data <- read_csv(simple_data_path,
                        col_types = 
                            cols(
  `Employee Name` = col_character(),
  `ID #` = col_character(),
  `Phone Number` = col_character(),
  Address = col_character(),
  Email = col_character(),
  Pledge = col_number(),
  Team = col_factor(),
  `Date Hired` = col_date(format = "%m/%d/%y")
)
)
```
```{r}
str(simple_data)
```


## Back to Superheroes

```{r}
heroes_path <- here("Datasets", "heroes_information.csv")
heroes <- read_csv(heroes_path, 
                   na = c("-", "-99"),
                   col_types = 
                       cols(
  X1 = col_double(),
  name = col_character(),
  Gender = col_factor(),
  `Eye color` = col_character(),
  Race = col_character(),
  `Hair color` = col_character(),
  Height = col_double(),
  Publisher = col_character(),
  `Skin color` = col_character(),
  Alignment = col_factor(),
  Weight = col_double()
)
                       )
```

Let's continue our EDA: Let's try another covariation question: How does gender relate to alignment?

One way would be to make a table:
```{r}
heroes %>%
    count(Gender, Alignment)
```

Longer way:
```{r}
heroes %>%
    filter(!is.na(Gender), !is.na(Alignment)) %>%
    group_by(Gender, Alignment) %>%
    summarize(n = n())
```

Or maybe we could use a bar chart:
```{r}
heroes %>%
    filter(!is.na(Gender), !is.na(Alignment)) %>%
    ggplot() + 
        geom_bar(aes(x = Alignment, fill = Gender), position = "dodge") +
        labs(title = "Alignment by Gender", y = NULL)
```

What about percentages?
```{r}
heroes %>%
    filter(!is.na(Gender), !is.na(Alignment)) %>%
    group_by(Gender, Alignment) %>%
    summarize(n = n()) %>%
    mutate(total = sum(n), percent = n / total * 100)
```

