---
title: "More-Purrr"
author: "John W. Hoggard"
date: "11/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

Let's find averages of elements in the `mpg` data set:
```{r}
glimpse(mpg)
```

```{r}
mpg %>%
    select(displ, year, cyl, cty, hwy) %>%
    map(mean)
```

Let's generate some random data with different means:
```{r}
list_of_means <- c(1, 5, 10, 15)
data_list <- vector(mode = "list", 4)
for ( i in seq_along(list_of_means) ) {
    data_list[[i]] <- rnorm(20, mean = list_of_means[[i]])
}
data_list
```

By default, it returns a list:
```{r}
map(data_list, mean) %>% as_vector
```

There is also a `map_dbl` which will give back a *vector*, if every result is really a double:
```{r}
map_dbl(data_list, mean)
```

What are the standard deviations?
```{r}
map_dbl(data_list, sd)
```

You can also do other functions, like `map_chr`, `map_lgl`, `map_int`, etc.

How else could we get 20 random normals with different means?
```{r}
map(
    c(1, 5, 10, 20),
    function(x) { rnorm(20, mean = x)}
)
```

We even have a `map_dfc` for output which is a data frame, created by columns:
```{r}
map_dfc(
    c(1, 5, 10, 20),
    function(x) { rnorm(20, mean = x)}
)
```





