---
title: "More Relational"
author: "John W. Hoggard"
date: "10/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(here)
library(tidyverse)
```
Grab some data from last time:
```{r}
dataL <- tribble(
    ~ID, ~LeftData,
    #---|----------
    "a", "a data",
    "b", "b data",
    "c", "c data",
    "d", "d data"
)

dataR <- tribble(
    ~ID, ~RightData,
    "a",  10,
    "b",  20,
    "d",  30,
    "e",  40
)
```


```{r}
tutor_logs <- read_csv(
    here("Datasets", "Tidying", "Tutoring.csv"),
    col_types = cols(
  Date = col_date(format = "%m/%d/%y"),
  Tutor = col_character(),
  Class = col_character(),
  `Time (in hours)` = col_double()
)
)

tutors <- read_csv(here("Datasets", "Tidying", "Tutors.csv"))
```

## Filtering Joins

Find out what is in or not in another table.

**Semi-join**:  We want only stuff in Table A which has a match in Table B:  
`semi_join(A, B)`.
```{r}
dataL
```
```{r}
dataR
```

```{r}
semi_join(dataL, dataR)
```

With the tutoring data:  Who in the tutoring log is actually in our personnel file?
```{r}
semi_join(tutor_logs, tutors)
```
Didnt' find a matching variable name!
```{r}
glimpse(tutor_logs)
```
```{r}
glimpse(tutors)
```
Specify that the "First" in tutors is linked with "Tutor" in tutor_logs:
```{r}
semi_join(tutor_logs, tutors, by = c("Tutor" = "First"))
```

**Anti-join**:  If we want to know what is in A that is *not* in B, we can use
`anti_join(A, B)`.
```{r}
anti_join(dataL, dataR, by = "ID")
```

You might use the same thing to "double-check" your filtering join for the tutors above:
```{r}
anti_join(tutor_logs, tutors, by = c("Tutor" = "First"))
```

Oh yeah--Bob is really Robert in the tutors file!

Could fix it:
```{r}
tutor_logs[tutor_logs$Tutor == "Bob", 2] <- "Robert"
```

Now my filtering join is more responsible:
```{r}
semi_join(tutor_logs, tutors, by = c("Tutor" = "First"))
```

## Multiple Matches

```{r}
dataL <- tribble(
    ~ID, ~LeftData,
    #---|----------
    "a", "a data",
    "a", "a2 data",
    "a", "a3 data",
    "b", "b data",
    "c", "c data",
    "d", "d data",
    "d", "d2 data"
)

dataR <- tribble(
    ~ID, ~RightData,
    "a",  10,
    "a",  11,
    "b",  20,
    "d",  30,
    "e",  40
)
```

An inner join chooses all common variables, one for each line from dataL and dataR:
```{r}
inner_join(dataL, dataR)
```

Left join is same deal, but only get IDs from the left table:
```{r}
left_join(dataL, dataR, by = "ID")
```

But a semi-join doesn't produce multiple matches; it just filters out what isn't in the second table:
```{r}
semi_join(dataL, dataR)
```

Absolutely *everything* is a full join:
```{r}
full_join(dataL, dataR)
```

## Set Operations

There are three operations: `intersect` (find what's in both), `union` (to find everything in either one), or `set_difference` (to find out what's in one but not the other).

```{r}
tutors2 <- read_csv(
    here("Datasets", "Tidying", "Tutors2.csv")
)
```

The columns have to match!
```{r}
union(tutors, tutors2)
```

What's in both?
```{r}
intersect(tutors, tutors2)
```

```{r}
tutors_temp <- tutors
tutors_temp[2, 3] <- "@0123"
tutors_temp
```

```{r}
union(tutors2, tutors_temp)
```

```{r}
intersect(tutors2, tutors_temp)
```

Then what's in tutors but not in tutors2?
```{r}
setdiff(tutors, tutors2)
```

Is it symmetric?
```{r}
setdiff(tutors2, tutors)
```





