---
title: "Strings"
author: "John W. Hoggard"
date: "10/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = T}
library(tidyverse)
library(here)
library(stringr)
```

## Basic Strings

Anything in quotes:
```{r}
"Hello, world!"
```

How do I include quotes?  Like "I said, "Hello, World!""?
```{r}
"I said \"Hello, World!\""
```
A backslash "escapes" the next character. If we want to see the string (rather than the escaped representation) there is a command `writeLines`:
```{r}
writeLines("I said \"Hello, World!\"")
```
Vectorized:
```{r}
writeLines(c("Hello", "I said \"Hello\""))
```
Other common escapes include newlines:
```{r}
writeLines("This is one line\nAnd this is another.")
```
```{r}
"el Ni\u0f1o"
```
Note above: the "\u0f1" represents a Unicode character.

Basic operations:  Length of a string via `str_length`:
```{r}
str_length("John Hoggard")
```
Base R equivalent:  `nchar`:
```{r}
nchar("John Hoggard")
```
The stringr version has a few nice features, like we can use it on factors: Recall in `iris` we had Species:
```{r}
str_length(iris$Species)
```

## Transformations in stringr

Example sentence:
```{r}
my_sent <- "the quick brown fox jumped over the lazy dog."
```

We can change cases for example:
```{r}
str_to_upper(my_sent)
```
```{r}
str_to_sentence(my_sent)
```

Some of this depends on *locale*.
```{r}
str_to_upper("i", locale = "tr")
```

We can eliminate extra whitespace with either str_trim (removes extra space at beginning or end of a string) or str_squish (removes all extra whitespace):
```{r}
x <- "              This string           has      lots of    whitespace.    "
```
```{r}
str_trim(x)
```
```{r}
str_squish(x)
```

## Combining Strings

To concatenate strings, we have `str_c`:
```{r}
str_c("Hello", "world", sep = " there ")
```

More interesting with vectors:
```{r}
str_c("var", 1:5, sep = "-")
```
You can also collapse together the resulting vector with a `collapse` argument:
```{r}
str_c("var", 1:5, collapse = ", ")
```

There are some similar functions in base R; see `paste` and `paste0`:
```{r}
paste("var", 1:5)
```

```{r}
paste0("var", 1:5)
```

Let's run an email scam:
```{r}
tutors <- read_csv(here("Datasets", "Tidying", "Tutors.csv"))
```

```{r}
str_c(
    "Congratulations ", 
    tutors$First, " ",
    tutors$Last, "!!!\n",
    "You may have already won fabulous prizes!"
) %>%
    writeLines()
```

## Subsetting

We can take a subset out of a string with `str_sub(<string>, <start>, <end>)`:
```{r}
my_name <- "John Hoggard"
str_sub(my_name, 1, 4)
```

```{r}
str_sub(my_name, -3, -1)
```

```{r}
str_sub(my_name, start = 1:5, end = -1)
```

```{r}
str_sub(my_name, 3, 5) <- "x"
my_name
```

(That is, we can actually *assign* into the result of the substring.)


