---
title: "Dates"
author: "John W. Hoggard"
date: "10/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
```

## Dates in Base R

Different formats in R for holding dates:  POSIXct and POSIXlt:
```{r}
(right_now <- Sys.time())
```

```{r}
unclass(as.POSIXct(right_now))
```

```{r}
unclass(as.POSIXlt(right_now))
```

Base R can convert strings to times with a format string, using `strptime`:
```{r}
(a_date <- strptime("10/27/2020", "%m/%d/%Y"))
```

```{r}
str(a_date)
```
I can actually extract pieces from the result:

```{r}
a_date$mday
```

## Conversions with `lubridate`

Have a family of functions for converting strings to dates:
```{r}
mdy(c("Jan 5, 1975", "5/5/1980", "6-12-2012")) 
```

```{r}
mdy("July 17, 1955")
```

Other orders exist also:
```{r}
ymd("2020-10-27")
```

We can also have date-time objects:
```{r}
ymd_hm("2010, March 11, 4:27 pm")
```

Can get even more specific, by including a timezone, with `tz` argument.  What's our timezone?
```{r}
Sys.timezone()
```

```{r}
ymd_hm("2010, March 11, 4:27 pm", tz = "America/New_York")
```

I can also subtract times:
```{r}
ymd_hm("2010, March 11, 4:27 pm", tz = "America/New_York") -
    ymd_hm("2010, March 11, 4:27 pm", tz = "GMT")
```

Other useful functions:
```{r}
today()
```


```{r}
now()
```

How long since the semester started?
```{r}
today() - ymd("2020-08-17")
```

More specific?
```{r}
now() - ymd_h("2020-08-17 8 am")
```

Find out how many days old you are today!
```{r}
today() - ymd(20000530)
```

Lubridate also includes some accessor functions to get pieces of a date/time object:
```{r}
year(a_date)
```
```{r}
month(a_date)
```
```{r}
hour(a_date)
```
```{r}
wday(a_date, label = TRUE)
```

What day of the week were you born on?
```{r}
wday(ymd(20000530), label = TRUE)
```

Can also produce date/time objects out of individual components with `make_datetime`:
```{r}
make_date(year = 2008, month = 11, day = 17)
```

Your accessor function can be used to change a date component too:
```{r}
a_date
```

```{r}
year(a_date) <- 2019
a_date
```

## Timespans

Differences are `difftime` type, specify with units.
```{r}
from_start <- now() - ymd_h(2020081708)
str(from_start)
```

A duration on the other hand is recorded as a number of seconds:
```{r}
str(as.duration(from_start))
```
```{r}
unclass(as.duration(from_start))
```

If I want a specific duration, we can this with a set of functions that start with d:
```{r}
# 10 seconds
dseconds(10)
```

```{r}
# 10 days
ddays(10)
```

What will the date be in 68 days?
```{r}
today() + ddays(68)
```
Apparently a number is interpreted as days:
```{r}
today() + 68
```

What will the date be in 68 hours:
```{r}
now() + dhours(68)
```

We can divide durations:  How much of this year has been spent in Fall semester?
```{r}
( as.duration( now() - ymd_h(2020081708))) / (as.duration(now() - ymd_h(2020010100)))
```

What percent of your life happened since you first started college?







