---
title: 'Day 5: DPLYR'
author: "John W. Hoggard"
date: "2020-09-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

More `dplyr`!

## Selecting Variables

Reminder:
```{r}
str(mpg)
```

Warm-up: Select the manufcaturer, model, city and highway mileage columns:
```{r}
head(
    select(mpg, c(manufacturer, model, cty, hwy))
)
```

Order matters:
```{r}
select(mpg, hwy, cty, manufacturer, model)
```

## Filtering Rows

Reminder: `iris`
```{r}
str(iris)
```

We use `filter` to pick which rows/observations to show
```{r}
filter(iris, Species == "virginica")
```

Note the `==`!  Use `==` to test for equality.

```{r}
filter(iris, Species = "virginica")
```

Nice error message in `dplyr`!

We can also pick anything with Sepal length of at least 5 cm:
```{r}
head(filter(iris, Sepal.Length >= 5))
```

```{r}
head(filter(iris, Sepal.Length > 5))
```

If we want two conditions, say Sepal length over 7 and sepal width over 3:
```{r}
filter(iris, Sepal.Length > 7, Sepal.Width > 3)
```

You can also combine with an and operation, abbreviated `&`:
```{r}
filter(iris, (Sepal.Length > 7) & (Sepal.Width > 3))
```

**Warning:**  There is also an `&&`, but don't use it on `filter` conditions:
```{r}
filter(iris, (Sepal.Length > 7) && (Sepal.Width > 3))
```

Or can be expressed with `|`:
```{r}
filter(iris, Sepal.Length > 7.5 | Sepal.Width > 4)
```

```{r}
str(mpg)
```


Check: Find the cars in `mpg` which are SUVs with highway mileage of at least 20mpg:
```{r}
filter(
    mpg, class == "suv", hwy >= 20
)
```

## Arrange the Rows

The `arrange` command will sort rows by a column:
```{r}
head(arrange(iris, Sepal.Length))
```

If we want descending order instead, there is a helper function, `desc()`:
```{r}
head(arrange(iris, desc(Sepal.Length)))
```

We can give multiple criteria to `arrange` as "tie-breakers":
```{r}
head(arrange(iris, desc(Sepal.Length), Petal.Length))
```

Can you find the shortest sepal length from the Virginica species? (You can use an intermediate variable if you want.)
```{r}
iris_virginica <- filter(iris, Species == "virginica")
arrange(iris_virginica, Sepal.Length)
```

```{r}
arrange(
    filter(iris, Species == "virginica"),
    Sepal.Length
)
```

## Using the Pipe

The pipe operator, `%>%`, lets us pass dataframes from one operation to the next.
```{r}
filter(iris, Species == "virginica") %>%
    arrange(Sepal.Length) %>%
    head()
```

In fact, I can just start with the data frame:
```{r}
iris %>%
    filter(Species == "virginica") %>%
    arrange(Sepal.Length) %>%
    head()
```

The SUVs with at least 20mpg in highway mileage, sorted by highway mileage:
```{r}
mpg %>%
    filter(class == "suv") %>%
    filter(hwy >= 20) %>%
    arrange(desc(hwy))
```

## Mutating the Data

We can add new variables based on old variables by mutating the data frame:
```{r}
# Convert Sepal lengths into inches:
mutate(iris,
       Sepal.Length.in = Sepal.Length / 2.54) %>%
    head()
```

```{r}
# Convert Sepal lengths into inches:
mutate(iris,
       Sepal.Length.in = Sepal.Length / 2.54,
       Sep.Len.in.plus1 = Sepal.Length.in + 1) %>%
    select(Sepal.Length, Sepal.Length.in, Sep.Len.in.plus1) %>%
    head()
```

Maybe we want ratio of Sepal.Length to Sepal.Width:
```{r}
mutate(iris,
       Ratio = Sepal.Length / Sepal.Width) %>%
    head()
```

## Summarizing

By itself, `summarize` is a little boring:
```{r}
summarize(iris,
          Median.Sep.Width = median(Sepal.Width),
          Mean.Sep.Width = mean(Sepal.Width),
          Sep.Width.SD = sd(Sepal.Width))
```

Can you find the median sepal length of the Setosa irises?
```{r}
iris %>%
    filter(Species == "setosa") %>%
    summarize(MedSepLength = median(Sepal.Length))
```

## Grouping Data

We can "group" the data together by some characteristic, then carry out summaries on the grouped data.

For example:
```{r}
iris_byspecies <- group_by(iris, Species)
head(iris_byspecies)
```

Now:
```{r}
summarize(iris_byspecies,
          mean(Sepal.Length), median(Sepal.Length), var(Sepal.Length))
```

There is a useful helper function `n()` which just counts how many there are:
```{r}
summarize(iris_byspecies,
          HowMany = n())
```

We could also:
```{r}
iris %>%
    group_by(Species) %>%
    summarize(mean(Sepal.Length), median(Sepal.Length))
```



