---
title: "Factors"
author: "John W. Hoggard"
date: "10/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Factors in Base R

We can create factors for categorical data in base R:
```{r}
eyecolor <- factor(
    c("brown", "blue", "brown", "brown", "hazel", "blue", "brown", "blue")
)
```

```{r}
str(eyecolor)
```

```{r}
eyecolor
```

What the "levels" are for your factor:
```{r}
levels(eyecolor)
```

By default alphabetical, but can change the order with a `levels` vector in the `factor` command:
```{r}
eyecolor <- factor(
    c("brown", "blue", "brown", "brown", "hazel", "blue", "brown", "blue"), 
    levels = c("hazel", "brown", "blue")
)
```

```{r}
levels(eyecolor)
```

By default, it doesn't matter for most things:
```{r}
eyecolor[1] < eyecolor[2]
```

Can create an *ordered* factor:
```{r}
eyecolor <- factor(
    c("brown", "blue", "brown", "brown", "hazel", "blue", "brown", "blue"), 
    levels = c("hazel", "brown", "blue"),
    ordered = TRUE
)
eyecolor[1] < eyecolor[2]
```
Can also adjust the levels directly:
```{r}
levels(eyecolor)[1] <- "Hazel"
```

```{r}
levels(eyecolor)
```

## With forcats: Lumping

Package forcats:
```{r}
library(forcats)
```

We can "lump together" your less frequently appearing categories into some sort of "other" category.  Use `fct_lump`, with argument `n` that tells us how many to keep.
```{r}
starwars %>%
    filter(!is.na(species)) %>%
    select(name, species) %>%
    mutate(species = fct_lump(species, n = 3)) %>%
    count(species) %>%
    arrange(desc(n))
```

Alternate form:  `lump_min` will group stuff with fewer than `min` entries
```{r}
starwars %>%
    filter(!is.na(species)) %>%
    select(name, species) %>%
    mutate(species = fct_lump_min(species, min = 2)) %>%
    count(species)
```
















