---
title: "Strings, Contd"
author: "John W. Hoggard"
date: "10/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## More on Using Regular Expressions

Let's say you wanted to take a vector and look for something like an a/A followed by any two letters:
```{r}
x <- c("Anything is a good thing", "aaaaaaa", "Something else here, good or bad.")
```

Usually we just put down a string:
```{r}
str_view_all(x, "[aA]..")
```

Really, the string is wrapped in a function called `regex`:
```{r}
str_view(x, regex("[aA].."))
```

You have some nice arguments.  For example, you can enable comments in regular expressions with `regex`:
```{r}
pattern <- regex(
    "[aA] # upper lower case A
    ..    # any two characters
    ",
    comment = TRUE
)
```
```{r}
str_view(x, pattern)
```

If you just want a fixed string, you can use `fixed` instead of `regex`:
```{r}
str_view("This is Mr. Smith", fixed("Smith"))
```

This can be problematic with other languages or special characters.

## Finding Matches

To find all the values in a vector which match a regular expression, we have `str_detect(<vector>, <regex>)`.
```{r}
x <- c("Abba", "Abracadabra", "oh no")
str_detect(x, "[Aa]..a")
```

Could use this in a filter, for example:
```{r}
x <- c("Alabama", "Arkansas", "National CS Data", "National Math Data")
str_detect(x, fixed("National"))
```

We could subset like so:
```{r}
x[str_detect(x, fixed("National"))]
```

But I can also use logical operations: ! (not), & (and), | (or):
```{r}
x[! str_detect(x, fixed("National"))]
```

There is a `words` vector with words in it. I could look for stuff that has a q, but not an r:
```{r}
pat1 <- str_detect(words, "q")
pat2 <- str_detect(words, "r")
words[ pat1 & (! pat2)]
```

```{r}
words[pat1 & pat2]
```

The `str_detect` is similar to `grepl` in base R, but the vector comes first.

## Counting Matches

We can use `str_count` to count matches:
```{r}
x <- c("Abba", "abracadabra", "aaaaaaaa", "oh no")
str_count(x, "[Aa]..a")
```

## Extraction

Sometimes we want to extract whatever matches from a string. Use `str_extract(vector, <pattern>)`:
```{r}
str_extract(x, "[Aa]..a")
```

If you want everything, you need `str_extract_all`:
```{r}
str_extract_all(x, "[Aa]..a")
```

You can get a so-called matrix if we add `simplify = TRUE`:

```{r}
str_extract_all(x, "[Aa]..a", simplify = TRUE)
```

Grades: Distributions of As, Bs, Cs, etc.?
```{r}
grades <- tribble(
    ~Name, ~Grade,
    "Bob", "B",
    "Cindy", "A-",
    "Doug", "B-H",
    "Erika", "AH",
    "Ginny", "D+",
    "Jazmine", "B+",
    "Frank", "D-",
    "Willy", "S"
    )
```

```{r}
grades %>%
    mutate(letter_grade = str_extract(Grade,
                                      "[ABCDFWUS]")) %>%
    count(letter_grade)
```

## Splitting Strings

We've already seen `separate` in `tidyr` that splits columns. (In fact, it will split on a regular expression, if you tell it to.)  If you just want to split strings, there is `str_split`:
```{r}
str_split("Now is the time for        all    good    people ... etc.",
          "\\s+")
```

## String Replacement

We can replace the found regular expressions with `str_replace` and `str_replace_all`:
```{r}
str_replace("Fred told me Hi", "Fred", "Velma")
```

```{r}
str_replace_all("`Ruh roh` said Scooby, really!", "[Rr]", "")
```

This can be useful trick for data cleaning:
```{r}
conditions <- c(
    "Snow", "rain", "clear", "Rainy", "Raining", "raining", "Raining", "Sunny",
    "rainy", "windy", "Clear", "Rain", "It was raining, bro"
)
str_replace_all(conditions, ".*[Rr]ain.*", "rain")
```





















