---
title: "Relational Data"
author: "John W. Hoggard"
date: "10/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(here)
library(tidyverse)
```

Start with some example tibbles:
```{r}
dataL <- tribble(
    ~ID, ~LeftData,
    #---|----------
    "a", "a data",
    "b", "b data",
    "c", "c data",
    "d", "d data"
)

dataR <- tribble(
    ~ID, ~RightData,
    "a",  10,
    "b",  20,
    "d",  30,
    "e",  40
)
```

```{r}
dataL
```

```{r}
dataR
```

## Joins

Putting the data together is called a *join*.

### Inner Join

Returns what is in each table, using a variable `by` to specify the (string) name of the variable to join by:

```{r}
inner_join(dataL, dataR, by = "ID")
```

If you leave out the `by` variable, it will attempt a "natural" join:
```{r}
inner_join(dataL, dataR)
```

### Outer Joins

We join up everything either on the left table (a left join) with everything that matches on the right, *or* join up everything on the right table (right join) with anything that matches in the left table:
```{r}
left_join(dataL, dataR, by = "ID")
```

```{r}
right_join(dataL, dataR, by = "ID")
```

### Full Join

Everything from both tables:
```{r}
full_join(dataL, dataR, by = "ID")
```

## Tutoring Data

Tutoring data:
```{r}
tutor_logs <- read_csv(
    here("Datasets", "Tidying", "Tutoring.csv"),
    col_types = cols(
  Date = col_date(format = "%m/%d/%y"),
  Tutor = col_character(),
  Class = col_character(),
  `Time (in hours)` = col_double()
)
)
```

We can use a left join to "look up" more information about the tutors, if we have another file with additional data. 

Grab extra data about the tutors:
```{r}
tutors <- read_csv(here("Datasets", "Tidying", "Tutors.csv"))
```

Now let's combine: Look up data for each tutor, and add it to the tutoring logs:
```{r}
left_join(tutor_logs, tutors, by = c("Tutor" = "First"))
```

Mostly good, except:  Oops, Bob is actually Robert.
```{r}
tutor_logs %>%
    mutate(Tutor = ifelse(
        Tutor == "Bob", 
        "Robert", 
        Tutor
    )) ->
    tutor_logs
```

```{r}
left_join(tutor_logs, tutors, by = c("Tutor" = "First")) ->
    tutor_data
```

Let's get the hours to send to HR to cut some paychecks:
```{r}
tutor_logs %>%
    group_by(Tutor) %>%
    summarize(Total = sum(`Time (in hours)`)) %>%
    left_join(tutors, by = c("Tutor" = "First")) %>%
    select(Tutor, Last, ID, Total) %>%
    rename(First = Tutor)
```

```{r}
orig_tutor_logs <- read_csv(
    here("Datasets", "Tidying", "Tutoring.csv"),
    col_types = cols(
  Date = col_date(format = "%m/%d/%y"),
  Tutor = col_character(),
  Class = col_character(),
  `Time (in hours)` = col_double()
)
)
```

If we had done an inner join instead:
```{r}
inner_join(orig_tutor_logs, tutors, by = c("Tutor" = "First"))
```

## AP Data

Data about the CS AP Exams in the US from 2006-2013:
```{r}
ap_data <- read_csv(
    na = c("*"),
    col_types = cols(
        .default = col_number(),
        state = col_character()
    ),
    here("Datasets", "pass_06_13.csv")
)
```

Now add some additional information: High School enrollments from
https://nces.ed.gov/programs/digest/d13/tables/dt13_203.20.asp:
Import this data:
```{r}
enrollment_data <- read_csv(
    here("Datasets", "HSEnrollment.csv")
)
```
Fix the NA rows:
```{r}
enrollment_data %>%
    filter(!is.na(State)) ->
    enrollment_data
```

Pivot this longer to get state and year we can look up:
```{r}
enrollment_data %>%
    pivot_longer(
        cols = -State,
        names_to = "year",
        values_to = "Enrollment",
        names_prefix = "Fall "
    ) %>%
    mutate(year = as.numeric(year)) ->
    long_enroll_data
```

Now let's join together `ap_data` and `long_enroll_data`:
```{r}
left_join(ap_data, long_enroll_data, 
          by = c("state" = "State", "year") ) ->
    full_ap_data
```

What percentage of students were taking the AP exam by state and year?
```{r}
full_ap_data %>%
    filter(!grepl("National", state)) %>% # strip out stuff that contains "National"
    mutate(Percent = total / Enrollment) %>%
    select(state, year, total, Enrollment, Percent)
```



