---
title: "Finish GGPlot and Start DPLYR"
author: "John W. Hoggard"
date: "2020-0902"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Saving a Plot

```{r}
library(ggplot2)
ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width)) + 
    geom_point(aes(color = Species)) +
    labs(title = "Sepal width vs. Length")
```

Save to a file:
```{r}
ggsave("IrisPlot.png")
```

## Coordinates

Diamonds data set:
```{r}
str(diamonds)
```

What different values of clarity do we have?
```{r}
ggplot(diamonds) +
    geom_bar(aes(x = clarity))
```

What if I want to split up by color?
```{r}
ggplot(diamonds) +
    geom_bar(aes(x = clarity, fill = color))
```

We can adjust the position:  Identity plots in front:
```{r}
ggplot(diamonds) +
    geom_bar(aes(x = clarity, fill = color), position = "identity")
```

```{r}
ggplot(diamonds) +
    geom_bar(aes(x = clarity, fill = color), position = "stack")
```

Or dodge to go side-by-side:
```{r}
ggplot(diamonds) +
    geom_bar(aes(x = clarity, fill = color), position = "dodge")
```

Can flip the coordinates:
```{r}
ggplot(diamonds) +
    geom_bar(aes(x = clarity, fill = color), position = "stack") +
    coord_flip()
```

But geom_bar computes counts.  If we just want data values, use geom_col:
```{r}
# DSCI 301 enrollment data
enrol_301 <- data.frame(
    year = c(2019, 2020),
    students = c(8, 10)
)
ggplot(enrol_301) + 
    geom_col(aes(x = year, y = students))
```

Let's go back to the diamonds:
```{r}
ggplot(diamonds) +
    geom_bar(aes(x = 1, fill = color)) + 
    coord_polar(theta = "y")
```

## Using DPLYR for Data Manipulation

```{r}
library(dplyr)
```

The `dplyr` package is for manipulating data frames.  The major techniques are:

    * Selecting columns (variables)
    * Filter rows (observations)
    * Arrange rows
    * Mutating variables (making/changing variables)
    * Summarizing data.
    
### Selecting Columns

```{r}
str(iris)
```

To select certain columns, use the `select` command:
```{r}
head(select(iris, Sepal.Length, Species))
```

We can also select ranges:
```{r}
head(select(iris, Petal.Length:Species))
```

Select everything *except* Species:
```{r}
head(
    select(iris, -Species)
)
```

Select everything except the length variables:
```{r}
head(
    select(iris, -c(Sepal.Length, Petal.Length))
)
```

In fact, we can be fancier:
```{r}
head(
    select(iris, -contains("Length"))
)
```










