---
title: "Functions"
author: "John W. Hoggard"
date: "11/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Functions

You know how to *use* functions:
```{r}
vec <- 1:10
mean(vec)
```
```{r}
sum(vec)
```
```{r}
length(vec)
```

We can also write our own functions with the `function`:
```{r}
f <- function(x) {
    x^2
}
```

I can use this just like a built-in function:
```{r}
f(-3)
```

By default, the last line evaluated in a function is the "return value", but you can be more explicit, with a return statement:
```{r}
f <- function(x) {
    2*x - 3  # Silly, because it does nothing
    return(x^2)
}
```

```{r}
f(2)
```

```{r}
g <- function(x) {
    x <- x + 1;
    x^2
}
```

So for example:
```{r}
g(2)
```

We can have multiple parameters:
```{r}
add_two <- function(x, y) {
    return(x + y)
}
```

```{r}
add_two(5, 3)
```

Of course, we can do things with vectors, too:
```{r}
f(1:10)
```

Like suppose we want a "deviation" function to get a typical deviation from the mean (not the *standard deviation*):
```{r}
deviation <- function(vec) {
    avg <- mean(vec)
    abs(vec - avg)
}
```

```{r}
deviation(1:5)
```

```{r}
x1 <- 1:5
mean(x1)
abs(x1 - mean(x1))
```

```{r}
library(dplyr)
```

```{r}
iris %>% mutate(
    dev_Sepal.Length = deviation(Sepal.Length)
)
```

Now we have an easy function to find the mean (`mean`), but let's write our own function to find the mean of a vector:
```{r}
my_mean <- function(vec) {
    total <- sum(vec)
    num_elts <- length(vec)
    total / num_elts
}
```

```{r}
my_mean(1:10)
```

```{r}
mean(1:10)
```

Built-in `mean` does have some extra features:
```{r}
x2 <- c(1, 2, 5, 7, NA)
mean(x2, na.rm = TRUE)
```

```{r}
my_mean(x2)
```

How about the standard deviation:
```{r}
my_sd <- function(vec) {
    mu <- mean(vec)
    sqrt(
        sum( (vec - mu)^2) / (length(vec) - 1)
        )
}
```


```{r}
my_sd(1:10)
```

```{r}
sd(1:10)
```

I can add extra arguments:
```{r}
my_sd <- function(vec, rem_na = FALSE) {
    mu <- mean(vec, na.rm = rem_na)
    no_na <- sum(is.na(vec))
    sqrt(
        sum( (vec - mu)^2, na.rm = rem_na) / (length(vec) - no_na - 1)
        )
}
```


```{r}
sd(x2, na.rm = TRUE)
```

```{r}
my_sd(x2, rem_na = TRUE)
```

```{r}
my_sd(x2, rem_na = FALSE)
```

Do have named arguments:
```{r}
my_sd(vec = x2, rem_na = TRUE)
```

```{r}
my_sd(x2, TRUE)
```

```{r}
sd(x = x2, na.rm = TRUE)
```

The names always allow you to specify the arguments:
```{r}
sd(na.rm = TRUE, x = x2)
```

Since we set `rem_na = FALSE` by default, we don't have to provide that argument:
```{r}
my_sd(x2)
```

```{r}
my_sd(x2, TRUE)
```

Functions don't actually have to be named:
```{r}
(function(x) {x+5})(2)
```

```{r}
integrate(function(x) x^2, 0, 2)
```

## Conditionals

```{r}
x <- 5
if (x < 5) {
    print("Less than 5.")
} else if (x > 5) {
    print("Greater than 5.")
} else {
    print("Equal to 5.")
}
```

Let's write a maximum function, to find the biggest of two values:
```{r}
my_max <- function(x, y) {
    if (x > y) {
        return(x)
    } else {
        return(y)
    }
}
```

```{r}
my_max(2, 3)
```

```{r}
my_max(3, 2)
```

```{r}
my_max(5, 5)
```

Or you can do it the CSCI 130/230/330 way:
```{r}
my_max <- function(x, y) {
    if (x > y) {
        maximum <- x
    } else {
        maximum <- y
    }
    return(maximum)
}
```

Let's make a summary stats function, which returns different stats depending on whether we want parametric or non-parametric.  In the case of parametric, let's get back a mean and a standard deviation.  In non-parametric, let's use median and mean average deviation.
```{r}
summary_stat <- function(x, parametric = TRUE) {
    if ( parametric ) {
        return(
            c("Mean" = mean(x), "StdDev" = sd(x))
        )
    } else {
        return(
            c("Median" = median(x) , "MAD" = mad(x))
        )
    }
}
```

```{r}
summary_stat(1:10)
```

```{r}
summary_stat(1:10, parametric = FALSE)
```













