---
title: "More EDA"
author: "John W. Hoggard"
date: "9/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read the superhero dataset:
```{r}
library(here)
library(tidyverse)
heroes_path <- here("Datasets", "heroes_information.csv")
heroes <- read_csv(heroes_path, 
                   na = c("-", "-99", "-99.0"),
                   col_types = 
                       cols(
  X1 = col_double(),
  name = col_character(),
  Gender = col_factor(),
  `Eye color` = col_character(),
  Race = col_character(),
  `Hair color` = col_character(),
  Height = col_double(),
  Publisher = col_character(),
  `Skin color` = col_character(),
  Alignment = col_factor(),
  Weight = col_double()
)
                       )
```

## Covariation: Two continuous variables

Weight and height are both numeric/continuous variables.

```{r}
heroes %>%
    ggplot(aes(x = Height, y = Weight)) +
    geom_point()
```

What's going on with the outliers?
```{r}
filter(heroes, Height < 100, Weight > 250)
```

Giganta:  She's been recorded as 62.5 *meters*, not cm.

```{r}
filter(heroes, Height > 500)
```

Reasonable ranges for heights/weights:
```{r}
heroes %>%
    filter(!is.na(Height), !is.na(Weight), Height < 500) %>%
    ggplot(aes(x = Height, y = Weight)) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE)
```

Alternatively: We can also discretize height, for example:
```{r}
heroes <- heroes %>%
    mutate(HeightRange = cut_number(Height, 10))
```

```{r}
heroes %>%
    filter(!is.na(Height), !is.na(Weight)) %>%
    ggplot(aes(x = HeightRange, y = Weight)) +
    geom_boxplot()
```

## AP Data

Load the data:
```{r}
ap_path <- here("Datasets", "pass_06_13.csv")
ap_data <- read_csv(ap_path, 
                    col_types = cols(
  state = col_character(),
  year = col_double(),
  schools_offering = col_double(),
  total = col_double(),
  passed = col_double(),
  female = col_double(),
  female_passed = col_number(),
  black = col_double(),
  black_passed = col_number(),
  black_male = col_double(),
  black_male_passed = col_number(),
  black_female = col_double(),
  black_female_passed = col_number(),
  hispanic = col_double(),
  hispanic_passed = col_number(),
  hispanic_female = col_double(),
  hispanic_female_passed = col_number(),
  hispanic_male = col_double(),
  hispanic_male_passed = col_double(),
  white = col_double(),
  white_passed = col_double(),
  white_male = col_double(),
  white_female = col_double(),
  asian = col_double(),
  asian_passed = col_double(),
  asian_male = col_double(),
  asian_female = col_double(),
  male = col_double(),
  male_passed = col_double()
))
```

All of them:
```{r}
spec(ap_data)
```

It appears all parsing failures are where there was *only* a `*` instead of a number.







