---
title: "Formatting"
author: "John W. Hoggard"
date: "11/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

Three things to improve RMarkdown:

  *  RMarkdown options
  *  Tables
  *  More ggplot options.
  
## RMarkdown Options

The "chunks" (the pieces which show up between the three single quotes).

  * `include = FALSE`:  Don't show either input or output from this command.
  * `echo = FALSE`: Don't print the command (but do print the output).
  * `message`, `warning`: control whether messages and warning are printed in the final file.
  
There's a ton of options for `knitr`: For more information, see
 https://yihui.org/knitr/.
 
(Also Chapter 21.)

## Tables

By default, tables that you print in RMarkdown show up as just "raw" output.
```{r}
head(mpg)
```

Alternative: You wrap these in a function `knitr::kable`:
```{r}
knitr::kable(
    head(mpg)
)
```

Other arguments can add more details:

  * `caption` to create a caption over the table.
  * `col.names`: specify strings to use for the column names.
  * `digits`: specifies how many decimal places to round to.
  
More in Chapter 22.

Markdown by itself is a little limited; to get fancier you can specify the kind of output (html, pdf) and then add more options with a package `kableExtra` (needs installation).
```{r}
knitr::kable(
    head(mtcars)
) %>%
    kableExtra::kable_styling(
        bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

## More GGPlot Options

(See stuff in Chapter 22.)

Non-data elements (like axes labels, etc) are adjusted as part of a "theme":
```{r}
mpg %>%
    ggplot(aes(x = displ, y = hwy, color = drv)) + 
    geom_point() +
    theme(axis.text.x = element_text(size = 12, # Size of text
            angle = 90,
            vjust = .5))
```

We can also change scaling, such as for exponential type data:
```{r}
exprands <- tibble(x = exp(rnorm(1000)))
ggplot(exprands, aes(x)) + 
    geom_histogram(bins = 30) + 
    scale_x_log10()
```

### Colors and Themes

You *can* specify colors yourself:
```{r}
ggplot(iris, aes(x = Sepal.Width, y = Sepal.Length, color = Species)) + 
    geom_point() +
    scale_color_manual(values = c("purple", "yellow", "black"))
```

If you want to know, you can find the color names with `colors`:
```{r}
colors()
```

Pre-picked options in `scale_color_brewer`:
```{r}
ggplot(iris, aes(Sepal.Width, Sepal.Length, color = Species)) +
    geom_point() + 
    scale_color_brewer(palette = "YlOrRd" )
```

There are other sets of color scales designed for continuous gradations.
```{r}
df <- tibble(x = rnorm(10000), y = rnorm(10000))
ggplot(df, aes(x, y)) + 
    geom_bin2d() + 
    coord_fixed() +
    scale_fill_binned(type ="viridis")
```

### Themes

We can add a theme:
```{r}
mpg %>%
    ggplot(aes(displ, hwy, color = drv)) + 
    geom_point() + 
    theme_grey()
```

There are more available in `ggthemes` package:
```{r}
mpg %>%
    ggplot(aes(displ, hwy, color = drv)) + 
    geom_point() +
    ggthemes::theme_wsj()
```

```{r}
mpg %>%
    ggplot(aes(displ, hwy, color = drv)) + 
    geom_point() +
    ggthemes::theme_fivethirtyeight()
```













