---
title: "Further Modeling"
author: "John W. Hoggard"
date: "11/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Linear Models

```{r}
library(tidyverse)
```

```{r}
glimpse(mtcars)
```

Built a model of mpg based on hp:
```{r}
hp_mod <- lm( mpg ~ hp, data = mtcars)
hp_mod
```

How good is the fit?
```{r}
library(modelr)
mtcars %>% add_predictions(hp_mod, var = "Pred") %>%
    add_residuals(hp_mod, var = "Resid") %>%
    ggplot(aes(x = hp, y = mpg)) +
    geom_point() +
    geom_line(aes(y = Pred))
```

Residuals?
```{r}
mtcars %>% 
    add_residuals(hp_mod, var = "Resid") %>%
    ggplot(aes(x = Resid)) +
    geom_histogram(bins = 10)
```

## Multiple Regressors

```{r}
toy2 <- tibble(
    x = 1:30,
    y = runif(30)*10,
    Out = 3*x -2*y + 5 + rnorm(30)
)
```

```{r}
mod_toy <- lm(Out ~ x + y, data = toy2)
mod_toy
```

Residuals?
```{r}
toy2 %>% add_residuals(mod_toy, var = "Resid") %>%
    ggplot(aes(x = Resid)) +
    geom_histogram(bins = 12)
```

"Average" residual?
```{r}
toy2 %>% add_residuals(mod_toy, var = "Resid") %>%
    summarize(   sqrt(sum(Resid^2)/n()) )
```

Note: We usually look at a residual *standard error*, which uses the number of data points minus the number of degrees of freedom in your model.
```{r}
toy2 %>% add_residuals(mod_toy, var = "Resid") %>%
    summarize(   sqrt(sum(Resid^2)/(n() - 3)) )
```

We can still look for patterns:
```{r}
toy2 %>% 
    add_residuals(mod_toy, var = "Resid") %>%
    ggplot(aes(x = Out, y = Resid)) +
    geom_point()
```

Looks pretty random.

## Regression with Factors

Regression using factors (as inputs) is a little special:
```{r}
my_mt <- mtcars %>%
    mutate(gear = factor(gear),
           cyl = factor(cyl))
mod_gear <- lm(mpg ~ gear, data = my_mt)
mod_gear
```

```{r}
my_mt %>%
    count(gear)
```

R has produced "dummy variables" which represent true/false categories of having 4 or 5 gears. Default is 3 gears (the first factor).

```{r}
my_mt %>% 
    group_by(gear) %>%
    summarize(mean(mpg))
```

Why 24.53 for 4 gears and we get coefficient of 8.427?
```{r}
24.5333 - 16.10667
```

In other words, the `gear4` coefficient is the *difference* between the default (3 gears) and the new category (4 gears) for average mpg.
















