---
title: "Tidy Data with tidyr"
author: "John W. Hoggard"
date: "10/8/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(here)
```

## Tidy Data

Tidy data:

1.  Every variable in its own column
2.  Every observation in its own row
3.  Each value in its own cell

Four (five?) examples (built-in to `tidyr`) of various formats (from WHO TB cases data):
```{r}
table1
```

```{r}
table2
```

```{r}
table3
```

```{r}
table4a
```

```{r}
table4b
```

```{r}
table5
```
## Updates

Major upgrade last year.

What installed version?
```{r}
packageDescription("tidyr")$Version
```
We want version >= 1.  Can update just `tidyr` using `install.packages` again, or can update *everything* old with `update.packages(ask = FALSE)` to auto-update everything. 

## Gathering / Pivoting Longer

Sometimes the column headers actually have data in them, like `table4a`:
```{r}
table4a
```

The year is a variable.  We will pivot to a longer table using `pivot_longer`.

`pivot_longer` takes a data frome plus three variables:

  * Vector of columns to turn into a new name/value pair
  * `names_to`: Specifies what to name the column for the old names
  * `values_to`: Name of the new column for the values

```{r}
table4a %>%
    pivot_longer(
        cols = `1999`:`2000`,
        names_to = "Year",
        values_to = "Cases"
    )
```

A sample data frame:
```{r}
test_data <- 
    tribble(
        ~Name,   ~val1,      ~val2,       ~val3,
        #-------/---------/------------/--------
        "Bob",   "BobVal1", "BobVal2",  "BobVal3",
        "Mia",   "MiaVal1", "MiaVal2",  "MiaVal3",
        "Ora",   "OraVal1", "OraVal2",  "OraVal3"
    )
test_data
```

If we pivot to a longer form:
```{r}
test_data %>%
    pivot_longer(
        cols = 2:4,
        names_to = "ValName",
        values_to = "Value"
    ) -> test_long
test_long
```

### New vs. Old

The older way used `gather`, and the variables were `key` (name of column for the keys), and `value` (name of new column for values), then any columns you wanted to convert:

When pivoting longer:

|  Old     |  New                | Meaning                                    |
|----------|---------------------|--------------------------------------------|
| `key`    | `names_to`          |  What to call the new column of old titles |
| `value`  | `values_to`         |  What to call the new column of values     |
| columns  | vector of cols      |  Which columns to convert                  |


```{r}
table4a %>%
    gather(
        key = "Year",
        value = "Cases",
        `1999`, `2000`
    )
```


## Spreading / Pivoting Wider

Conversely, we may have a "key" column which we would like to split up into multiple other columns.  Can do this with a "reverse" operation, `pivot_wider`:  Have to specify `names_from` (which column will have the names of the new columns), and `values_from` (which column to grab values from).

```{r}
test_long %>%
    pivot_wider(
        names_from = ValName,
        values_from = Value
    )
```


### Old and New

Aside:  The old version of this was called `spread`, and required `key` and `value` columns:

When pivoting wider:

|  Old     |  New                | Meaning                                    |
|----------|---------------------|--------------------------------------------|
| `key`    | `names_from`        |  Column containing new col titles          |
| `value`  | `values_from`       |  Column containing values to fill with     |

```{r}
test_long %>%
    spread(
        key = ValName,
        value = Value
    )
```

### Examples

Gradebook: May have recorded data individually as it comes in, with something like Name, Assignment, Grade, etc...
```{r}
grades <- read_csv(
    here("Datasets", "Tidying", "Grades.csv")
)
```
```{r}
glimpse(grades)
```

Would be nice to have columns for things like Quiz 1, HWK 2, etc:
```{r}
grades %>%
    pivot_wider(
        names_from = Assignment,
        values_from = Grade,
    )
```

Let's fix the bad variable name:
```{r}
grades %>%
    mutate(Assignment = ifelse(Assignment == "test 1",
                                "Test 1",
                                Assignment)) -> grades
```

Try again:
```{r}
grades %>%
    pivot_wider(
        names_from = Assignment,
        values_from = Grade,
    )
```

What's up with Sarah's Homework 3?
```{r}
filter(grades, Name == "Sarah Studious")
```





