---
title: "Exploratory Data Analysis"
author: "John W. Hoggard"
date: "9/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Goal of Exploratory Data Analsis: Understand the data and find patterns.

Common questions:

  * What values are present?
  * Variation within a variable?
  * Are there outliers?
  * Covariation?
 
Superhero data set from Kaggle: https://www.kaggle.com/ishajain30/super-heroes/data

Use package `here`:
```{r}
library(here)
```

Can get the path easily:
```{r}
hero_file <- here("Datasets", "heroes_information.csv")
heroes <- read.csv(hero_file, header = TRUE)
library(tidyverse)
```

## Categorical Variables

What are the eye colors?
```{r}
heroes %>%
    group_by(Eye.color) %>%
    summarize(count = n())
```

Shortcut:  `count`:
```{r}
heroes %>%
    filter(Eye.color != "-") %>%
    count(Eye.color) %>%
    arrange(desc(n)) %>%
    filter(n > 5)
```

## Numerical Variable

Let's look at height. What's the distribution?
```{r}
heroes %>%
    ggplot() +
    geom_histogram(aes(x = Height))
```

We could restrict the x-coordinates a bit with `coord_cartesian(xlim = ...)`:
```{r}
heroes %>%
    ggplot() +
    geom_histogram(aes(x = Height)) +
    coord_cartesian(xlim = c(-150, 0))
```

What's the chunk near -100?
```{r}
filter(heroes, Height < -90) %>%
    select(name, Race, Height)
```

I think -99 is used for NA in Height column. Let's make this explicit:
```{r}
heroes <- heroes %>%
    mutate(
        Height = ifelse( Height == -99,
            NA,
            Height
        )
    ) 
```

Back to histogram:
```{r}
heroes %>%
    filter(!is.na(Height)) %>%
    ggplot() + 
        geom_histogram(aes( x = Height ), binwidth = 10) +
    coord_cartesian(xlim = c(0, 300))
```

What are these numbers actually? Let's look up one:
```{r}
heroes %>%
    filter(name == "Iron Man") %>%
    select(name, Height)
```

These appear to be in cm.  How big is 500 cm?  `r 500 / 2.54 / 12` feet.

Who is over 500 cm?
```{r}
heroes %>%
    filter(Height > 500)
```

Who is small?
```{r}
heroes %>%
    filter(Height < 100) %>%
    select(name, Height) %>%
    arrange(desc(Height))
```

I am pretty sure *most* of these are in cm, but some *are not*.

We could also produce some summary statistics:
```{r}
heroes %>%
    summarize(
        Median = median(Height, na.rm = T),
        Mean = mean(Height, na.rm = T),
        SD = sd(Height, na.rm = T)
    )
```

## Covariation: Height vs. Race

How does height vary with race?
```{r}
heroes %>%
    filter(Race != "-") %>%
    group_by(Race) %>%
    summarize(AvgHeight = mean(Height, na.rm = T), count = n()) %>%
    filter(count >= 10) %>%
    arrange(desc(AvgHeight))
```

Distributions by race?
```{r}
heroes %>%
    filter(Race != "-") %>%
    group_by(Race) %>%
    filter(n() > 5) %>%
    ungroup() %>%
    filter(!is.na(Height)) %>%
    ggplot() +
        geom_boxplot(aes(x = Race, y = Height)) + 
        coord_flip()
```














