---
title: "More Iteration"
author: "John W. Hoggard"
date: "11/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## A Bit More on Preallocation

You can preallocate a list:
```{r}
ten_different_samples <- vector(mode = "list", 10)
means <- 1:10
for (index in seq_along(means)) {
    ten_different_samples[[index]] <- rnorm(20, mean = means[[index]])
}
```

```{r}
ten_different_samples
```

Similar:
```{r}
x <- c("Bob", "Sally", "Francine", "Arthur")
out <- ""
for (name in x) {
    out <- paste(out, name)
}
out
```

More efficient:
```{r}
paste(x, collapse = " ")
```

You *could* add rows to a dataframe using `rbind`:
```{r}
rbind(head(iris), c(1.2, 2.1, 3.4, 1.6, "virginica"))
```

But doing this repeatedly is expensive. Don't loop through adding more stuff onto the end of a data frame.  Instead, if we want to put lots of stuff together, use `dplyr::bind_rows`:
```{r}
dplyr::bind_rows(iris[1:2,], iris[15:20,], iris[22:24,])
```

## While Loop

Needed in particular if we don't know how many times to repeat something:
```{r, eval = F}
while (condition) {
    <statements>
}
```

How long does it take to get something less than 0 out of a random normal distribution with mean of 1 and standard deviation of 1?
```{r}
count <- 1
x <- rnorm(1, mean = 1)
while ( x >= 0 ) {
    print(paste(x, ">= 0; try #", count))
    count <- count + 1
    x <- rnorm(1, mean = 1)
}

```
```{r}
x
```

```{r}
count
```

## Maps with purrr

Part of the tidyverse (`purrr`) for iteration.  Uses a `map` function to "map" a function over a vector or list.
```{r, eval = F}
map(<list>, <f>)
```

will apply `f` to each element of `list`, and return the resulting list.

```{r}
library(purrr)
map(ten_different_samples, mean)
```

Returns a list, but we could reformat this:
```{r}
map(ten_different_samples, mean) %>% as_vector()
```








