---
title: "More Factors with forcats"
author: "John W. Hoggard"
date: "10/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
```

## Factors: Reordering

We have some helper functions to reorder:

  * `fct_inorder`: Reorders a factor variable by the first appearance of a factor.
  * `fct_infreq`: Reorders a factor variable by the frequency of a factor (most to least common)
  * `fct_inseq`: If it's numeric, put it in numerical order.

```{r}
eyecolor <- factor(
    c("brown", "blue", "hazel", "brown", "brown", "brown", "blue"),
    ordered = TRUE)
eyecolor
```

```{r}
fct_inorder(eyecolor)
```

```{r}
fct_infreq(eyecolor)
```

Why? Star Wars data:
```{r}
starwars %>%
    filter(!is.na(species)) %>%
    mutate(species = fct_infreq(fct_lump(species, n = 8))) %>%
    count(species) %>%
    ggplot(aes(x = species, y = n)) +
        geom_col()
```

Let's do similar with the heros data set:
```{r}
heros <- read_csv(na = c("-99.0", "-"),
    here("Datasets", "heroes_information.csv"), 
    col_types = cols(
  X1 = col_double(),
  name = col_character(),
  Gender = col_factor(),
  `Eye color` = col_factor(),
  Race = col_factor(),
  `Hair color` = col_factor(),
  Height = col_double(),
  Publisher = col_character(),
  `Skin color` = col_factor(),
  Alignment = col_factor(),
  Weight = col_double()
)
)
```

Let's look at the eye colors:
```{r}
heros %>%
    filter(!is.na(`Eye color`)) %>%
    mutate(eye_color = fct_infreq(fct_lump(`Eye color`, n = 8))) %>%
    count(eye_color) %>% arrange(desc(n)) %>%
    ggplot(aes(x = eye_color, y = n)) + 
        geom_col()
```

We can also reorder factors by a second vector using `fct_reorder`. By default, it reorders by the median value taken from the second vector.
```{r}
data <- tibble(
    ec = eyecolor, 
    n = c(5, 4, 3, 6, 5, 6, 1)
)
data
```
```{r}
fct_reorder(data$ec, data$n)
```

For example:
```{r}
iris %>%
    ggplot(aes(x = fct_reorder(Species, Sepal.Width),
                                y = Sepal.Width)) + 
        geom_boxplot()
```

## Changing Factors

We can manually change factors:
```{r}
eyecolor
```

```{r}
levels(eyecolor)[3] <- "Hazel"
eyecolor
```

But from forcats, we have `fct_recode`, which is simpler to understand: We specify what to name the new categories:
```{r}
fct_recode(eyecolor,
           "Brown" = "brown",
           "Blue" = "blue")
```

We also have `fct_collapse`, which does something similar, but you get to provide a vector of old factors on the right hand side.
```{r}
fct_collapse(eyecolor,
             "Brown/Hazel" = c("brown", "Hazel"))
```

In the heros set:
```{r}
heros %>%
    filter(!is.na(`Eye color`)) %>% 
    count(`Eye color`)
```

There are a certain number of "mixed" eye colors; why don't we put those together?
```{r}
heros %>%
    filter(!is.na(`Eye color`)) %>%
    .$`Eye color` -> eyecols
```

```{r}
mixed <- str_subset(eyecols, "/")
```

Now collapse:
```{r}
fct_collapse(heros$`Eye color`, 
             "Mixed" = mixed) %>% levels
```

Also have a `fct_reorder2`, which reorders the factor based on the values in *two* vectors, x and y, according to the largest y value for the last x value.  (What????)

Let's go back to our AP data set:
```{r}
ap_data <- read_csv(here("Datasets", "pass_06_13.csv"),
                    na = c("", "*"),
                    col_types = cols(
                        .default = col_number(),
                        state = col_character()
                    ))
```

Now: Let's look at number passed per year, by various categories:
```{r}
ap_data %>%
    filter(year > 2005) %>%
    filter(str_detect(state, "National.*CS")) %>%
    pivot_longer(cols = ends_with("_passed"),
                 names_to = "Category",
                 values_to = "Passed") %>%
    mutate(Category = fct_reorder2(Category, year, Passed)) %>%
    ggplot(aes(x = year, y = Passed, color = Category)) +
        geom_line() + geom_point()
```

It makes the line graph match the legend (on the right).














