---
title: "Tidy Tuesday"
author: "John W. Hoggard"
date: "9/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Tidy Tuesday

Download the package with `install.packages("tidytuesdayR")`, and load the data set:
```{r}
library(tidyverse)
tuesdata <- tidytuesdayR::tt_load('2020-09-22')
```
```{r}
str(tuesdata)
```

Look at, say, expeditions:
```{r}
str(tuesdata$expeditions)
```

Each is available to extract:
```{r}
tuesdata$expeditions %>% head
```

## Expeditions

```{r}
exped <- tuesdata$expeditions
str(exped)
```

Expeditions by year?
```{r}
count(exped, year) %>%
    ggplot(aes(x = year, y = n)) +
    geom_line()
```

Seems like the number of expeditions has increased by year for some time, with a recent drop--or are there just fewer recorded on the most recent? Says through Spring 2019, so 2019 is probably incomplete-others are maybe OK?
```{r}
filter(exped, year > 2010) %>% count(year)
```
```{r}
exped %>%
    filter(year < 2019) %>%
    count(year) %>%
    ggplot(aes(x = year, y = n)) +
    geom_line()
```

Have deaths been changing over the years?
```{r}
exped %>%
    group_by(year) %>%
    summarize(Deaths = sum(member_deaths)) %>%
    ggplot(aes(x = year, y = Deaths)) + 
    geom_line()
```

Has risen some, although not I think in proportion to total increase in number of members:
```{r}
exped %>%
    group_by(year) %>%
    summarize(Deaths = sum(member_deaths), TotalMembers = sum(members)) %>%
    mutate(PercDeaths = Deaths / TotalMembers) %>%
    ggplot(aes(x = year, y = PercDeaths)) + 
    geom_line()
```

Year when everyone died?
```{r}
exped %>%
    group_by(year) %>%
    summarize(Deaths = sum(member_deaths), TotalMembers = sum(members)) %>%
    mutate(PercDeaths = Deaths / TotalMembers) %>%
    filter(PercDeaths == 1)
```
```{r}
filter(exped, year == 1934)
```

